---
title: "Lab 1 - feedback solutions"
author: "Ines Arana"
date: "Session 2023-24"
output:
  html_document: default
  pdf_document: default
---

## Libraries

Loading the required library. In the lab sheet this is done on page 5.
Note: ggplot2 is used for plotting data. As this library is loaded as part of tidyverse, if the latter is loaded, there is no need to explicitly load the former.

```{r warning=FALSE, message=FALSE}
library(tidyverse)
```

## Variables 

```{r}
x <- 7
x
```

### Vectors

More complex variables

```{r}
energy_source <- c("oil","gas","wood","sun","wind","wave",
                   "hydroelectric","nuclear")
energy_source
```

Displaying the 3rd element

```{r}
energy_source[3]
```

Displaying items 4 to 6

```{r}
energy_source[4:6]
```

Creating a new vector from values in another vector

```{r}
traditional_source <- c(energy_source[1:3],energy_source[5],
                        energy_source[7])

traditional_source
```


Getting the number of items in a vector.

```{r}
length(traditional_source)
```

## Arithmetic operations

Creating some variables using operators.

```{r}
x <- 4
y <- x + 20
z <- y/x
```

Checking the values of x, y and z

```{r}
x
y
z
```

### Operations with a vector

```{r}
numbers <- c(1,2,3,4,5,6,7,8,9,10)
numbers * 10
```

Using 2 vectors

```{r}
numbers * c(10,20,30)
numbers * c(10,20,30,10,20,30,10,20,30,10)
```

## Logical operators

```{r}
numbers > 7
```

Combination of expressions - numbers greater than 7 or smaller than 2

```{r}
(numbers > 7) | (numbers< 2)

```

## Various examples of plots using the power of a number dataset.

First we create the dataset with numbers 1 to 6.

```{r}
my_nums <-(1:6)
powers <- my_nums^2
```

Using a tibble

```{r}
powerData <- tibble(my_nums,powers)
powerData
```

Using a data frame

```{r}
powerdata2<- data.frame(my_nums,powers)
powerdata2
```

We can plot now. This code will not do much, as there is no specification of how to plot the data.

```{r}
ggplot(data=powerData)
```

We can add to the above code to state we want a scatterplot (use geom_point) and which the x and y variables are.


```{r} 
ggplot(data=powerData) + 
  geom_point(aes(x=my_nums, y=powers))
```


The code above is the same as the one below, where each line contains one aspect of the definition of the plot.

```{r}
p <- ggplot(data=powerData)
p<- p + geom_point(aes(x=my_nums, y=powers))
p
```

Adding a title, labels and a subtitle.

```{r}
p <- p + labs(title= "Numbers against their powers")
p <- p + labs( x = "Number", y = "Power of number")
p <- p + labs(subtitle = "Simple scatterplot")
p
```

All the code put together, plus adding colour to the points.

```{r}
p <- ggplot(data=powerData)
p <- p + geom_point(aes(x=my_nums, y=powers), colour="red")
p <- p + labs(title= "Numbers against their powers")
p <- p + labs( x = "Number", y = "Power of number")
p <- p + labs(subtitle = "Simple scatterplot")
p
```



# Exercise 1


Anscombe Quartet exercise

Plotting each dataset.

### Dataset 1
```{r}
set1_x <- c(10,8,13,9,11,14,6,4,12,7,5)
set1_y <- c(8.04,6.95,7.58,8.81,8.33,9.96,7.24,4.26, 10.84,4.82,5.68)
data1 <- tibble(set1_x,set1_y)
ggplot(data=data1) + geom_point(aes(x=set1_x, y=set1_y))
```



### Dataset 2

```{r}

set2_x <- c(10,8,13,9,11,14,6,4,12,7,5)
set2_y <- c(9.14,8.14,8.74,8.77,9.26,8.10,6.13,3.10,9.13,7.26,4.74)
data2 <- tibble(set2_x,set2_y)
ggplot(data=data2) + geom_point(aes(x=set2_x, y=set2_y))
```


### Dataset 3

```{r}
set3_x <- c(10,8,13,9,11,14,6,4,12,7,5)
set3_y <- c(7.46,6.77,12.74,7.11,7.81,8.84,6.08,5.39,8.15,6.42,5.73)
data3 <- tibble(set3_x,set3_y)
ggplot(data=data3) + geom_point(aes(x=set3_x, y=set3_y))
```



### Dataset 4

```{r}
set4_x <- c(8,8,8,8,8,8,8,19,8,8,8)
set4_y <- c(6.58,5.76,7.71,8.84,8.47,7.04,5.25,12.50,5.56,7.91,6.89)
data4 <- tibble(set4_x,set4_y)
ggplot(data=data4) + geom_point(aes(x=set4_x, y=set4_y))
```




# Exercise 2



Including nice labels in the plots. We have already defined the data, so we just need to plot it


### Dataset 1
```{r}
p <- ggplot(data=data1)
p <- p + geom_point(aes(x=set1_x, y=set1_y))
p <- p + labs(title = "Anscombe Quartet", subtitle="Dataset 1", y= "y values", x = "x values") 
p

```


### Dataset 2
```{r}
p <- ggplot(data=data2)
p <- p + geom_point(aes(x=set2_x, y=set2_y))
p <- p + labs(title= "Anscombe Quartet", subtitle="Dataset 2", y= "y values", x = "x values")
p
```


### Dataset 3

```{r}
p <- ggplot(data=data3)
p <- p + geom_point(aes(x=set3_x, y=set3_y))
p <- p + labs(title = "Anscombe Quartet", subtitle="Dataset 3", y= "y values", x = "x values")
p
```



### Dataset 4

```{r}
p <- ggplot(data=data4)
p <- p + geom_point(aes(x=set4_x, y=set4_y))
p <- p + labs(title = "Anscombe Quartet", subtitle="Dataset 4", y= "y values", x = "x values")
p

```


# Exercise 3

Adding colour to the plots in previous exercise.


### Dataset 1
```{r}
p <- ggplot(data=data1)
p <- p + geom_point(aes(x=set1_x, y=set1_y), colour="purple")
p <- p + labs(title = "Anscombe Quartet", subtitle="Dataset 1", y= "y values", x = "x values") 
p

```


### Dataset 2
```{r}
p <- ggplot(data=data2)
p <- p + geom_point(aes(x=set2_x, y=set2_y), colour="orange")
p <- p + labs(title= "Anscombe Quartet", subtitle="Dataset 2", y= "y values", x = "x values")
p
```


### Dataset 3

```{r}
p <- ggplot(data=data3)
p <- p + geom_point(aes(x=set3_x, y=set3_y), colour = "blue")
p <- p + labs(title = "Anscombe Quartet", subtitle="Dataset 3", y= "y values", x = "x values")
p
```



### Dataset 4

```{r}
p <- ggplot(data=data4)
p <- p + geom_point(aes(x=set4_x, y=set4_y), colour="limegreen")
p <- p + labs(title = "Anscombe Quartet", subtitle="Dataset 4", y= "y values", x = "x values")
p

```



