---
title: "Lab 2 - feedback solutions"
author: "IA"
date: "Session 2023/24"
output:
  html_document: default
  pdf_document: default
---


```{r setupE}
knitr::opts_chunk$set(echo = TRUE)
```


Loading libraries. Note that tidyverse includes ggplot2, which is the library we need for plotting data.

```{r  include=FALSE}
library(tidyverse)

```


### Histogram

Histogram with the pressure1 dataset, displaying the pressure. 
First load the dataset, then produce the histogram.

Note: if you have loaded the data using the "Import Dataset" facility the  R variable containing the dataset is likely to be called "pressure1", so use this name below instead of "pressure1Data.

```{r}
m <- c("CMM004", "CMM007", "CMM020", "CMM022", "CMM024", "CMM028", "CMM012","CMM513")

mn <- c("SPE", "Intranet Systems Dev", "Data Vis and Analysis", "Databases", "OOP", "ITI and Administration", "Project Investigation", "MSc Project")

cr <- c(15,15,15,15,15,15, 15, 45)

whatToStudy <- data.frame(module = m, moduleName=mn, 
                          credits = cr, stringsAsFactors=T)
```


If we rather have a tibble we can use the tibble function - note that the vectors m, mn and cr were created above.

```{r}
whatToStudyT <- tibble(module = m, moduleName=mn,                           credits = cr)

whatToStudyT <- whatToStudyT  |> mutate(across(where(is.character),as_factor))
```

Accessing the 2nd column by name

```{r}
 
whatToStudy$moduleName
```

Accessing the 2nd column by index.

```{r}

whatToStudy[,2]
```

Accessing the 4th row.

```{r}
whatToStudy[4,]
```

Accessing the 3rd element in the moduleName column.

```{r}
whatToStudy$moduleName[3]
```

Accessing the 3rd element in the 2nd column.

```{r}
whatToStudy[3,2]
```

Getting the dimensions

```{r}
dim(whatToStudy)
```

Getting the number of rows
```{r}
nrow(whatToStudy)
```

Getting the number of columns
```{r}
ncol(whatToStudy)
```

```{r}
summary(whatToStudy)
```


## Exercise 1

Accessing the 2nd column by index.

```{r}

whatToStudyT[,2]
```

Accessing the 4th row.

```{r}
whatToStudyT[4,]
```

Accessing the 3rd element in the moduleName column.

```{r}
whatToStudyT$moduleName[3]
```

Accessing the 3rd element in the 2nd column.

```{r}
whatToStudyT[3,2]
```

Getting the dimensions

```{r}
dim(whatToStudyT)
```

Getting the number of rows
```{r}
nrow(whatToStudyT)
```

Getting the number of columns
```{r}
ncol(whatToStudyT)
```

```{r}
summary(whatToStudyT)
```

## Loading the pressureTemp dataset.

Loading the data onto a data frame

```{r}
pressureTemp <- read.csv("pressureTemp.csv", header=T, stringsAsFactors =T)

summary(pressureTemp)
```



Alternatively, load the data onto a tibble.
```{r}
pressureTempTib <- read_csv("pressureTemp.csv" )
pressureTempTib <- pressureTempTib |> 
			mutate(across(where(is.character),as_factor))

```

# Plots

## Histogram for pressure

```{r}
p <- ggplot(data = pressureTemp, aes(x=Pressure))
p <- p + geom_histogram()
p
```

Colouring it blue.
 
```{r}
p <- ggplot(data = pressureTemp, aes(x=Pressure))
p <- p + geom_histogram(fill="blue")
p
```

Setting the bin width to 4.
 
```{r}
p <- ggplot(data = pressureTemp, aes(x=Pressure))
p <- p + geom_histogram(binwidth=8, fill="blue")
p
```

Specifying that 10 bins are required (no binwidth in this case!).

```{r}
p <- ggplot(data = pressureTemp, aes(x=Pressure))
p <- p + geom_histogram(bins=10, fill="blue")
p
```

## Exercise 2

```{r}
p <- ggplot(data = pressureTemp, aes(x=Temperature))
p <- p + geom_histogram()
p


```

## Exercise 3

```{r}
p <- ggplot(data = pressureTemp, aes(x=Temperature))
p <- p + geom_histogram(fill="orange")
p


```

## Exercise 4

```{r}
p <- ggplot(data = pressureTemp, aes(x=Temperature))
p <- p + geom_histogram(bins=7, fill="lightblue")
p


```

## Exercise 5

```{r}
p <- ggplot(data = pressureTemp, aes(x=Temperature))
p <- p + geom_histogram(binwidth=2, fill="violet")
p


```

## Exercise 6

```{r}
p <- ggplot(data = pressureTemp, aes(x=Readby))
p <- p + geom_bar( fill="red")
p
```

## Exercise 7

```{r}
p <- ggplot(data = pressureTemp, aes(x=Readby))
p <- p + geom_bar( fill="brown")
p
```

## Reading the consumption dataset

```{r}
consumption <- read.csv("consumption.csv", header=T, stringsAsFactors = T)
```

## Column plot  for Energy in year X2035.


```{r}

p <- ggplot(data = consumption, aes(x=Energy, y=X2035)) 
p <- p + geom_col(fill="powderblue")
p
```

## Exercise 8

```{r}

p <- ggplot(consumption, aes(x=Energy, y=X2035)) 
p <- p + geom_col(fill="pink")
p
```

## Exercise 9

Loading the data file.

```{r}
consumptionTransposed <- read.csv("consumptionTransposed.csv",
                                    header=T, stringsAsFactors = T)

```


## Exercise 10

```{r}
p <- ggplot(consumptionTransposed, aes(x=Year, y=Oil)) 
p <- p + geom_col(fill="pink")
p
```


The year values are taken as numeric, so the space between columns varies.

## Exercise 11

Changing the year column from type numeric to type factor.

```{r}
consumptionTransposed$Year <- as.factor(consumptionTransposed$Year)

p <- ggplot(consumptionTransposed, aes(x=Year, y=Oil)) 
p <- p + geom_col(fill="pink")
p
```



## Exercise 12

```{r}
consumptionTransposed$Year <- as.factor(consumptionTransposed$Year)
```

## Exercise 13


```{r}
p <- ggplot(consumptionTransposed, aes(x=Year, y=Coal)) 
p <- p + geom_col(fill="red")
p
```

## Boxplots

```{r}
p <- ggplot(data= pressureTemp, aes(x=Readby, y=Temperature)) 
p <- p+ geom_boxplot()
p
```

## Exercise 14

```{r}
p <- ggplot(pressureTemp, aes(x=Readby, y=Temperature)) 
p <- p+ geom_boxplot(fill="red")
p
```

## Exercise 15

```{r}
p <- ggplot(pressureTemp, aes(x=Readby, y=Temperature)) 
p <- p+ geom_boxplot(fill=c("pink", "blue", "green"))
p
```


## Exercise 16

```{r}
p <- ggplot(pressureTemp, aes(x=Readby, y=Temperature)) 
p <- p+ geom_boxplot(fill=terrain.colors(3))
p
```



```{r}
p <- ggplot(pressureTemp, aes(x=Readby, y=Temperature)) 
p <- p+ geom_boxplot(fill=topo.colors(3))
p

```


## Exercise 17

```{r}
p <- ggplot(pressureTemp, aes(x=Readby, y=Pressure)) 
p <- p+ geom_boxplot(fill=topo.colors(3))
p

```


