---
title: "Lab3"
author: '007'
date: "2024-02-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load library tidyverse
```{r}
library(tidyverse)
```

#laod ggplot
```{r}
library(ggplot2)
```

#download data-set consumptionNoOil
```{r}
consumptionNoOil <- read.csv("consumptionNoOil.csv", header = T, stringsAsFactors = T)
```

```{r}
summary(consumptionNoOil)
#head(consumptionNoOil, 10)

```

#Aesthetics
```{r}
cno <- ggplot(data = consumptionNoOil, aes(x=MTOE, y = EstMTOE))
cno <- cno + geom_point()
cno
```
#can also be written as
```{r}
rm(cno)
cno <- ggplot(consumptionNoOil, aes(MTOE, EstMTOE))
cno <- cno + geom_point(colour="darkgreen")
cno

```

#using column source and shape
```{r}
rm(cno)
cno <- ggplot(consumptionNoOil, aes(MTOE, EstMTOE, colour=Source, shape = Source))
cno <- cno + geom_point()
cno
```

#using line plot to display time data
```{r}
rm(cno)
cno <- ggplot(consumptionNoOil, aes(Year, MTOE, colour=Source))
cno <- cno + geom_line()
cno
```

#Groups plot
#case 1: Multiple groups with one aesthetics
#plot year vs consumption by energy source
```{r}
rm(cno)
cno <- ggplot(consumptionNoOil, aes(Year, MTOE, group=Source))
cno <- cno + geom_line(size = 2, colour="lightblue",)
cno
```

#using point plot
```{r}
rm(cno)
cno <- ggplot(consumptionNoOil, aes(Year, MTOE, group=Source))
cno <- cno + geom_point(colour="lightpink")
cno <- cno + geom_line(colour="lightblue")
cno
```

#adding shape diagram
```{r}
rm(cno)
cno <- ggplot(consumptionNoOil, aes(Year, MTOE, group=Source))
cno <- cno + geom_point(shape = 23, colour="red", size = 3,fill = "lightblue")
#cno <- cno + geom_line(colour="lightblue")
cno
```

#case 2: Different groups on different layers:
```{r}
rm(cno)
cno <- ggplot(consumptionNoOil, aes(Year, MTOE, group = Source))
cno <- cno + geom_line()
cno <- cno + geom_smooth(aes(group = 1), formula = 'y~x', method = "lm", size = 2, se = F)
cno
```

testing
```{r}
rm(cno)
cno <- ggplot(consumptionNoOil, aes(Year, MTOE, group = Source))
cno <- cno + geom_line()
cno <- cno + geom_smooth(aes(group = Source), formula = 'y~x',  method = "lm", size = 1, se = F)
cno
```

#download data-set ProvenOilReservesZerosT
```{r}
ProvenOilReservesZerosT <- read.csv("ProvenOilReservesZerosT.csv", header = T)
```

#replacing empty value(-) with NA(not available)
```{r}
ProvenOilReservesZerosT |> mutate_all(funs(replace(., .== "-", as.double(NA))))
```

#exercises
##download data-set ProvenOilReservesZerosT
```{r}
rm(ProvenOilReservesZerosT)
ProvenOilReservesZerosT <- read.csv("ProvenOilReservesZerosT.csv", header = T, stringsAsFactors = T)
```

#replacing empty value(-) with NA(not available)
```{r}
ProvenOilReservesZerosT |>  mutate_all(funs(replace(., .== "-", as.double(NA))))
```

#scatterplot year vs denmark
```{r}
prs <- ggplot(ProvenOilReservesZerosT, aes(MTBarrels, Denmark))
prs <- prs + geom_point(shape = 22, size = 4, colour = "darkblue" )
prs

```

#line-plot year vs italy
```{r}
rm(prs)
prs <- ggplot(ProvenOilReservesZerosT, aes(MTBarrels, Italy))
prs <- prs + geom_line(size = 2, colour = "orange" )
prs

```

#scatter-plot year vs united kingdom
```{r}
rm(prs)
prs <- ggplot(ProvenOilReservesZerosT, aes(MTBarrels, United.Kingdom,colour = United.Kingdom ))
prs <- prs + geom_point()
prs

```

#scatter-plot year vs Norway
```{r}
rm(prs)
prs <- ggplot(ProvenOilReservesZerosT, aes(MTBarrels, Norway,size=Norway ))
prs <- prs + geom_point(shape = 24, colour="red",)
prs
```

#download data-set ProvenOilReserveWEurope
```{r}
ProvenOilReserveWEurope <- read.csv("ProvenOilReserveWEurope.csv", header = T, stringsAsFactors = T)
```

#Line-plot Year by country
```{r}
rm(pre)
pre <- ggplot(ProvenOilReserveWEurope,aes(Year, MT.Barrels, colour = Country))
pre <- pre + geom_line()
pre
```

#Line-plot Year by country + summary smooth
```{r}
rm(pre)
pre <- ggplot(ProvenOilReserveWEurope,aes(Year, MT.Barrels, colour = Country))
pre <- pre + geom_line()
pre <- pre + geom_smooth(aes(group = 1), formula = 'y~x', method = "lm", size = 2, se = F)
pre
```

#using point(scattered-plot) Year by country + summary smooth
```{r}
rm(pre)
pre <- ggplot(ProvenOilReserveWEurope,aes(Year, MT.Barrels, colour = Country))
pre <- pre + geom_point()
pre <- pre + geom_smooth(aes(group = 1), formula = 'y~x', method = "auto",se = F)
pre
```

#selecting specific data to store in a variable
```{r}
to1990 <- ProvenOilReserveWEurope |> filter(Year %in% c(1980:1990))
summary(to1990)
```

#Heat Map
```{r}
Hmp <- ggplot(to1990, aes(Year, Country))
Hmp <- Hmp + geom_tile(aes(alpha=MT.Barrels), fill="blue")
Hmp
```

#Heat Map 1990 to 2001
```{r}
to2001 <- ProvenOilReserveWEurope |> filter(Year %in% c(1990:2001))
summary(to2001)
```

#plotting the Heatmap
```{r}
rm(Hmp)
Hmp <- ggplot(to2001, aes(Year, Country))
Hmp <- Hmp + geom_tile(aes(alpha=MT.Barrels), fill="red")
Hmp
```

