---
title: "Lab4"
author: '007'
date: "2024-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load library tidyverse
```{r}
library(tidyverse)
```

#laod ggplot
```{r}
library(ggplot2)
```

#load all datasets
```{r}
projectTeam <- read.csv("projectTeam.csv", header = T, stringsAsFactors = T)
importsExportsAmerica <- read.csv("importsExportsAmerica.csv", header = T, stringsAsFactors = T)
channels <- read.csv("channels.csv", header = T, stringsAsFactors = T)
universities <- read.csv("universities.csv", header = T, stringsAsFactors = T)


```

#trying out text geometry
```{r}
iea <- ggplot(importsExportsAmerica, aes(crude.imports, crude.exports, label=country))
iea <- iea + geom_text( color = "red")
iea
```

#trying geom_area
```{r}
rm(iea)
iea <- ggplot(importsExportsAmerica, aes(crude.imports, crude.exports, label=country))
iea <- iea + geom_area(color = "green", fill ="yellow")
iea
```

#trying out geom_polygon
```{r}
rm(iea)
iea <- ggplot(importsExportsAmerica, aes(crude.imports, crude.exports, label=country))
iea <- iea + geom_polygon(color="cyan", fill="pink")
iea
```

#trying out geom_tile()
```{r}

rm(iea)
iea <- ggplot(importsExportsAmerica, aes(crude.imports, crude.exports, label=country))
iea <- iea + geom_tile(color="blue")
iea
```

#trying out stacking for projectteam
```{r}
pjt <- ggplot(projectTeam, aes(site, fill=experience))
pjt <- pjt + geom_bar(position = "stack")
pjt
```

#trying out dodging
```{r}
rm(pjt)
pjt <- ggplot(projectTeam, aes(site, fill=experience))
pjt <- pjt + geom_bar(position = "dodge")
pjt
```

#trying out filling
```{r}
rm(pjt)
pjt <- ggplot(projectTeam, aes(site, fill=experience))
pjt <- pjt + geom_bar(position = "fill")
pjt
```

#trying out identity
```{r}
rm(pjt)
pjt <- ggplot(projectTeam, aes(site, fill=experience))
pjt <- pjt + geom_bar(position = "identity")
pjt
```
#not a usefull for bar plot
#doesn't show all data


#adjusting x and y axix of a geom_point
```{r}
rm(cno)
cno <- ggplot(consumptionNoOil, aes(Year,MTOE, colour=Source))
cno <- cno + xlim(2010,2035)
cno <- cno + geom_point()
cno
```

#adding titles, subtitles and labels to plots
```{r}
rm(cno)
cno <- ggplot(consumptionNoOil, aes(Year,MTOE, colour=Source))
cno <- cno + geom_point()
cno <- cno + labs(x="Year", y="Millions of tons of oil equivalent", title = "Energy consumption", subtitle = "Oil is excluded", colour="Energy source")
cno

```

#changing colours
```{r}
rm(cno)
colours <- c(coal="black", gas="yellow", renewables="green")
cno <- ggplot(consumptionNoOil, aes(Year,MTOE, colour=Source))
cno <- cno + geom_point()
cno <- cno + labs(x="Year", y="Millions of tons of oil equivalent", title = "Energy consumption", subtitle = "Oil is excluded", colour="Energy source")
cno <- cno + scale_colour_manual(values = colours)
cno
```

#adding lines to plot
```{r}
rm(cno)
cno <- ggplot(consumptionNoOil, aes(Year,MTOE, colour=Source))
cno <- cno + geom_point()
cno <- cno + geom_vline(aes(xintercept = 2015), linetype = "longdash", colour="green")
cno
```

#for horizontal line
```{r}
rm(cno)
cno <- ggplot(consumptionNoOil, aes(Year,MTOE))
cno <- cno + geom_point()
cno <- cno + geom_hline(aes(yintercept = 100), linetype = "dotdash", colour="green")
cno
```

#to add both vert and hrz
```{r}
rm(cno)
cno <- ggplot(consumptionNoOil, aes(Year,MTOE, colour=Source))
cno <- cno + geom_point()
cno <- cno + geom_hline(aes(yintercept = 100), linetype = "dotdash", colour="darkgreen")
cno <- cno + geom_vline(aes(xintercept = 2015), linetype = "twodash", colour="blue")
cno
```

#plotting pie chart
#counting number of staff per site
```{r}
rm(pieData)
pieData <- projectTeam |> count(site, name = "numberStaff")
names(pieData) <- c("site", "numberStaff")
# order data according to the site (important for placing labels)
pieData <- arrange(pieData, desc(site))
# create new column with position for label
pieData <- pieData |> mutate(positionLabel = cumsum(numberStaff) - 0.5 * numberStaff)

```

#creating the pie plot
```{r}
rm(pjt)
pjt <- ggplot(pieData, aes(x="", y= numberStaff, fill=site))
pjt <- pjt + geom_bar(width = 1, stat = "identity")
pjt <- pjt + coord_polar("y", start = 0)
# adding text labels
pjt <- pjt + geom_text(aes(y = positionLabel, label = numberStaff))
# Changing the colour scheme
pjt <- pjt + scale_fill_manual(values = c("red", "blue", "lightblue"))
# removing outer "ring"
pjt <- pjt + theme_void()
pjt
```

#yearly vs denman
```{r}
rm(prs)
prs <- ggplot(ProvenOilReservesZerosT, aes(MTBarrels, Denmark))
prs <- prs + geom_point(shape = 20, size = 4, colour = "darkblue" )
prs <- prs + labs(x="Year", y="Million tons oil equivalent", title = "Proven Oil Reserve in Denmark")
prs <- prs + geom_vline(aes(xintercept = MTBarrels), linetype = "longdash", colour="green")
prs
```

#yearly vs italy
```{r}
rm(prs)
prs <- ggplot(ProvenOilReservesZerosT, aes(MTBarrels, Italy))
prs <- prs + geom_line(size = 2, colour = "orange" )
prs <- prs + labs(x="Year", y="Million tons oil equivalent", title = "Proven Oil Reserve in Italy")
prs <- prs + geom_hline(aes(yintercept = 0.7), linetype = "longdash", colour="green")
prs
```

#yearly vs united kingdom
```{r}
rm(prs)
prs <- ggplot(ProvenOilReservesZerosT, aes(MTBarrels, United.Kingdom, colour= United.Kingdom))
prs <- prs + geom_point( )
prs <- prs + labs(x="Year", y="Million tons oil equivalent", title = "Proven Oil Reserve in United Kingdom", colour = "Size of Reserves")
prs
```

#pre year vs MTOB by country
```{r}
rm(pre)
pre <- ggplot(ProvenOilReserveWEurope, aes(Year, MT.Barrels, group = Country))
pre <- pre +geom_line(aes(colour=factor(Country)))
pre <- pre + labs(x="Year", y="Million tons Barrels", title = "Proven Oil Reserve in Western Europe", colour = "country")
pre
```

#plottng channels datset
```{r}
chl <- ggplot(channels, aes(length, area, size=n))
chl <- chl + geom_point(colour="pink")
chl <- chl + labs(x="length", y="area" , title="Perception - length vs. area", size= "Size of stimuli")
chl
```

#plotting above to go from 0 - 5
```{r}
rm(chl)
chl <- ggplot(channels, aes(length, area, size=n))
chl <- chl + geom_point(colour="pink")
chl <- chl + labs(x="length", y="area" , title="Perception - length vs. area", size= "Size of stimuli")
chl <- chl + ylim(0,5)
chl

```

#plotting universities
```{r}
rm(uni)
uni <- ggplot(universities,aes(status, fill=school))
#uni <- uni + geom_bar(position = "identity")
uni <- uni + geom_bar(position = "dodge")
#uni <- uni + geom_bar(position = "fill")
#uni <- uni + geom_bar(position = "stack")
#uni <- uni + geom_bar(position = "jitter")
uni
```

#ex11 pie chaet
```{r}
rm(uni)
universities2 <- universities |> filter(status %in% c("student", "both"))

# Counting the number of rows per school.
uniPieData <- universities2 |> count(school, name="numberStudents")

#Ordering data according to the school name (important for placing labels later on)
uniPieData <- arrange(uniPieData,desc(school))

# Creating a new column called positionLabel with position for label
uniPieData <-  mutate(uniPieData, positionLabel = cumsum(uniPieData$numberStudents) - 0.5*uniPieData$numberStudents)

# creating pie chart
uni <- ggplot(uniPieData,  aes(x="", y= numberStudents, fill = school))
uni <- uni + geom_bar(width = 1, stat="identity")
uni <- uni + coord_polar("y", start=0)

# Adding text labels
uni <- uni + geom_text(aes(y = positionLabel, label = numberStudents))

# Changing the colour scheme
uni <- uni +   scale_fill_manual(values = c("wheat", "lightblue", "cornflowerblue", "blue1")) 
# removing outer "ring"
uni <- uni + theme_void()
uni
```

