---
title: "lab5"
author: '007'
date: "2024-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load tidyverse
```{r}
library(tidyverse)
```


#load libraryies
```{r}
library(treemapify)
```

```{r}
library(ggplot2)
```

```{r}
library(ggraph)
```

```{r}
library(igraph)
```

```{r}
library(ggwordcloud)
```

#load dataset
```{r}
consumptionBySource <- read.csv("consumptionBySource.csv", header = T, stringsAsFactors = T)
arrivals <- read.csv("arrivals.csv", header = T, stringsAsFactors = T)
scottishSales <- read.csv("scottishSales.csv", header = T, stringsAsFactors = T)

```

#plotting several variables
```{r}
rm(chl)
chl <- ggplot(channels, aes(n))
chl <- chl + geom_line(aes(y = electric, colour="electric current"))
chl <- chl + geom_line(aes(y = length, colour="object's length"))
chl <- chl + geom_line(aes(y = area, colour= "object's area"))
chl <- chl + labs(x="Size of stimuli", y="Amount of sensation",title = "Different channels sensation according to level of stimuli",colour="channels")
chl <- chl + ylim(0,50)
chl
```

#Removing electrict colunm
#do not run
```{r}
rm(chl)
#channels$electric <- NULL
chl <- ggplot(channels, aes(n))
chl <- chl + geom_line(aes(y = length, colour="object's length"))
chl <- chl + geom_line(aes(y = area, colour= "object's area"))
chl <- chl + labs(x="Size of stimuli", y="Amount of sensation",title = "Different channels sensation according to level of stimuli",colour="channels")
chl
```

#restricting rows
#comming back to fix
```{r}
library(tidyverse)
#rm(chl)
head(channels)
channels2 <- channels |> filter(electric < 50)

```

#Facet
```{r}
rm(pre)
pre <- ggplot(ProvenOilReserveWEurope, aes(Year))
pre <- pre + geom_line(aes(y = MT.Barrels, colour=Country))
pre <- pre + facet_grid(Country ~ .)
pre <- pre + labs(title="Proven oil reserves in Western Europe", y = "Billion barrels of oil equivalent")
pre
```

#making RStudio decide which y scale is best for each country
```{r}
rm(pre)
pre <- ggplot(ProvenOilReserveWEurope, aes(Year))
pre <- pre + geom_line(aes(y = MT.Barrels, colour=Country))
pre <- pre + facet_grid(Country ~ ., scales = "free_y")
pre <- pre + labs(title="Proven oil reserves in Western Europe", y = "Billion barrels of oil equivalent")
pre
```

#using linetype to diffrentiate countries
```{r}
rm(pre)
pre <- ggplot(ProvenOilReserveWEurope, aes(Year))
pre <- pre + geom_line(aes(y = MT.Barrels, colour=Country, linetype=Country ))
pre <- pre + facet_grid(Country ~ ., scales = "free")
pre <- pre + labs(title="Proven oil reserves in Western Europe", y = "Billion barrels of oil equivalent", colour = "European country", linetype ="European country")
pre

```

#using point in plotting same graph
```{r}
rm(pre)
pre <- ggplot(ProvenOilReserveWEurope, aes(Year))
pre <- pre + geom_point(aes(y = MT.Barrels, colour=Country,))
pre <- pre + facet_grid(Country ~ ., scales = "free_y")
pre <- pre + labs(title="Proven oil reserves in Western Europe", y = "Billion barrels of oil equivalent", colour = "European country",)
pre
```

#using both type in plotting same graph
```{r}
rm(pre)
pre <- ggplot(ProvenOilReserveWEurope, aes(Year))
pre <- pre + geom_point(aes(y = MT.Barrels, colour=Country,))
pre <- pre + geom_line(aes(y = MT.Barrels, colour=Country,))
pre <- pre + facet_grid(Country ~ ., scales = "free", space = "free_y")
pre <- pre + labs(title="Proven oil reserves in Western Europe", y = "Billion barrels of oil equivalent", colour = "European country",)
pre
```

#tree MAaps
```{r}
rm(sts)
sts <- ggplot(scottishSales, aes(area = Sales, fill = Category, subgroup=Category, label=City))
sts <- sts + geom_treemap()
sts <- sts + geom_treemap_subgroup_border()
sts <- sts + geom_treemap_text()
sts <- sts + geom_treemap_subgroup_text()
sts <- sts + labs(title = "Amount of sale by to product category and city", colour= "Sale category")
sts
```

#adding more parameteres to center text
```{r}
rm(sts)
sts <- ggplot(scottishSales, aes(area = Sales, fill = Category, subgroup=Category, label=paste(City,Sales,sep ="\n")))
sts <- sts + geom_treemap()
sts <- sts + geom_treemap_subgroup_border(colour = "white", size = 5)
sts <- sts + geom_treemap_text(place = "centre", size = 11, grow =TRUE)
sts <- sts + geom_treemap_subgroup_text(place = "centre", grow = TRUE, alpha = 0.25, colour="black", fontface ="italic")
sts <- sts + labs(title = "Amount of sale by to product category and city", colour= "Sale category")
sts
```

#Dendrograms
#starting by creating a hierarchy
```{r}
link_top_2 <- scottishSales |> select(Concept, City) |> unique() |> rename(from=Concept, to=City)
link2_3 <- scottishSales |> select(City, Sales) |> unique() |> rename(from=City, to=Sales)
link2_3$to <- as.factor(link2_3$to)
links <- rbind(link_top_2, link2_3)
dendro <- graph_from_data_frame(links)
```

#ploting the dendrogram
```{r}
rm(sts)
sts <- ggraph(dendro, layout = 'dendrogram', circular = FALSE)
sts <-sts + geom_node_label(aes(label= name), size=3, repel = TRUE, colour="darkgreen", fill="lightyellow", show.legend = TRUE)
sts <- sts + geom_edge_diagonal(colour="darkgreen")
sts <- sts + geom_node_point(colour="darkgreen")
sts <- sts + labs(title="Sales in various stores ")
sts
```

#plotting Wordclouds
```{r}
destinations <- as.data.frame(table(arrivals$Destination))
#giving the columns proper names
names(destinations) <- c("destination","count")
m <- mean(destinations$count)
destinations <- destinations |> filter(count>m)

```

#producing the wordcloud
```{r}
rm(arv)
set.seed(123)
arv <- ggplot(destinations, aes(label = destination, size=count))
arv <- arv + geom_text_wordcloud_area(colour="blue")
arv <- arv + scale_size_area(max_size = 40)
arv <- arv + labs(title = "Most common destination ports for vessels departing from Barcelona")
arv
```

