---
title: "LAB6-1"
author: '007'
date: "2024-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load required dataset
```{r}
browserdata <- read.csv("browserdata.csv", header = T, stringsAsFactors = T)
```

#load required dataset
```{r}
consumptionTransposed2 <- read.csv("consumptionTransposed2.csv", header = T, stringsAsFactors = T)
```

#Applying different functions to view 
```{r}
#summary(consumptionTransposed2)
#str(consumptionTransposed2)
#class(consumptionTransposed2)
#head(consumptionTransposed2)
tail(consumptionTransposed2)
```

#obtaining summary of a column
```{r}
consumptionTransposed2 |> select(Oil) |> summary()
```

#using the pull() function to extract data
```{r}
consumptionTransposed2 |> pull(Gas) |> mean()
consumptionTransposed2 |> pull(Gas) |> quantile()
consumptionTransposed2 |> pull(Gas) |> fivenum()
consumptionTransposed2 |> pull(Gas) |> sd()
```

#summarizing browserdata 
```{r}
summary(browserdata)
str(browserdata)
class(browserdata)
head(browserdata)
tail(browserdata)
```
#using the pull function for browser data
```{r}
browserdata |> pull(level) |> mean()
browserdata |> pull(level) |> median()
browserdata |> pull(level) |> sd()
browserdata |> pull(level) |> var()
browserdata |> pull(level) |> range()
browserdata |> pull(level) |> IQR()
browserdata |> pull(level) |> min()
browserdata |> pull(level) |> max()
browserdata |> pull(level) |> quantile()
browserdata |> pull(level) |> fivenum()
```

#frequency count
```{r}
counts <- browserdata |> pull(Browser) |> table()
counts
```
#to get proportions
```{r}
prop.table(counts)
```

#getting a two way table
```{r}
freqBrowser <- xtabs(~ Browser+Membership, data = browserdata)
freqBrowser
```

#to get probabilities of table
```{r}
prop.table(freqBrowser, 1) #each row add up to 1
prop.table(freqBrowser, 2) #each column add up to 1
```

#coverting todata frame(using freqBrowser)
```{r}
e <- as.data.frame(freqBrowser)

```

#plot when analysing single variables
```{r}
#for a simple bar plot
freqb <- ggplot(e, aes(x= Membership, y=Freq, fill=Browser))
freqb <- freqb + geom_bar(stat="identity")
freqb <- freqb + labs(title = "Browser use by membership type", x = "Membership type", y = "Frequency count") #adding description
freqb <- freqb + scale_fill_brewer(palette = "Set1") #changing colour scheme and background
freqb <- freqb + theme_classic()
freqb

```

#showing the above data in a different column
```{r}
freqb <- ggplot(e, aes(x= Membership, y=Freq, fill=Browser))
freqb <- freqb + geom_bar(stat="identity", position = position_dodge())
freqb <- freqb + labs(title = "Browser use by membership type", x = "Membership type", y = "Frequency count") #adding description
freqb <- freqb + scale_fill_brewer(palette="Dark2") #changing colour scheme and background
freqb <- freqb + theme_classic()
freqb
```

#plotting histogram for daily time spent
#also adding colour, titles, labels and line
```{r}
rm(mpd)
mpd <- ggplot(browserdata, aes(x= MinsPerDay))
mpd <-  mpd + geom_histogram(aes(y=..density..), col="red", fill="red", binwidth = 20)
mpd <- mpd + labs(x="Average time spent online (mins/day)", y="Dendity")
mpd <- mpd + geom_density(col="blue")
mpd <- mpd + theme_classic()
mpd
```

#using Dot plots
```{r}
brd <- ggplot(browserdata, aes(x=Use))
brd <- brd + geom_dotplot(col="red", fill="red", binwidth = 0.05)
#add labels
brd <- brd + labs(x="Use level", y="")
brd <- brd + theme_classic()
brd
```

#using box plot
```{r}

```

